Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-05-30 manuscript/manuscript.Rmd
use PLOS latex template

---
title: "Comparing human and model-based forecasts of COVID-19 in Germany and Poland"
header-includes:
    - \usepackage{lineno}
    - \linenumbers
    -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
linestretch: 2 
bibliography:
  - references.bib
  - GermanPolishpaper.bib
csl: vancouver.csl
---

*Nikos I. Bosse, Sam Abbott, Johannes Bracher, Habakuk Hain, Billy J. Quilty, Mark Jit, Centre for the Mathematical Modelling of Infectious Diseases COVID-19 Working Group, Edwin van Leeuwen, Anne Cori, Sebastian Funk*

@@

# Abstract

Forecasts based on epidemiological modelling have played an important role in shaping public policy throughout the COVID-19 pandemic. This modelling combines knowledge about infectious disease dynamics with the subjective opinion of the researcher who develops and refines the model and often also adjusts model outputs. Developing a forecast model is difficult, resource- and time-consuming. It is therefore worth asking what modelling is able to add beyond the subjective opinion of the researcher alone. To investigate this, we analysed different real-time forecasts of cases of and deaths from COVID-19 in Germany and Poland over a 1-4 week horizon submitted to the German and Polish Forecast Hub. We compared crowd forecasts elicited from researchers and volunteers, against a) forecasts from two semi-mechanistic models based on common epidemiological assumptions and b) the ensemble of all other models submitted to the Forecast Hub. We found crowd forecasts, despite being overconfident, to outperform all other methods across all forecast horizons when forecasting cases (weighted interval score relative to the Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational models performed comparably better when predicting deaths (rel. WIS 1.26), suggesting that epidemiological modelling and human judgement can complement each other in important ways.

# Author summary

Mathematical models of COVID-19 have played a key role in informing governments across the world. While mathematical models are informed by our knowledge of infectious disease dynamics, they are ultimately developed and iteratively adjusted by the researchers and shaped by their subjective opinions. To investigate what modelling is able to add beyond the subjective opinion of the researcher alone, we compared human forecasts with model-based predictions of COVID-19 cases and deaths submitted to the so-called German/Polish Forecast Hub (which collates a variety of models from a range of teams).  
We found that our human forecasts consistently outperformed an aggregate of all available model-based forecasts when predicting cases, but not when predicting deaths. Our findings suggest that human insight may be most valuable when forecasting highly uncertain quantities, which depend on many factors that are hard to model using equations, while mathematical models may be most useful in settings like predicting deaths, where leading indicators with a clear connection to the target variable are available. This potentially has very relevant policy implications, as agencies informing policy-makers could benefit from routinely eliciting human forecasts in addition to model-based predictions to inform policies.

@@

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish Forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control [@ecdcDownloadHistoricalData2020]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@rkiRKICoronavirusSARSCoV22021]. These data are subject to reporting artefacts, (such as for example delayed case reporting in Poland on the 24th November, @forsal.plRozbieznosciStatystykachKoronawirusa2020), changes in reporting over time, and variation in testing regimes (for example in Germany from the 11th of November on, @aerzteblattSARSCoV2DiagnostikRKIPasst2020). The ECDC data as well as the data published by the Polish Ministry of Health were also subject to data revisions, although most of them (with a notable exception of a data update for October 12 2020 in Germany) only affected daily, not weekly data (see Figures \@ref(fig:daily-truth-update) and \@ref(fig:weekly-truth-update)).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the `shiny` and `golem` R packages [@golem; @shiny] and is available in the `crowdforecastr` R package [@crowdforecastr]. To make a forecast in the application participants could select a predictive distribution (with the default being log-normal) to represent the probability that the forecasted quantity took certain values. Median and width of the uncertainty could be adjusted by either interacting with a figure showing their forecast or providing numerical values (see screenshot in Figure \@ref(fig:screenshot) in the SI). The default shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e. as $\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )$). A comparison of the crowd forecasts against the default baseline shown in the application is displayed in Figure \@ref(fig:compare-forecasters) in the Appendix. 
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from  @ourworldindataCOVID19DataExplorer2020. These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time. From November 26 2020 on we displayed weekly small reports with a visualisation of past forecasts and scores on our website, epiforecasts.io.

@@

The study was approved by the Observational / Interventions Research Ethics Committee at the London School of Hygiene & Tropical Medicine, LSHTM Ethics Reference: 22290.

@@

For the main analysis we focused mostly on two week ahead forecasts, as COVID-19 forecasts, especially for cases, were in the past found to have poor predictive performance  beyond this horizon [@bracherShorttermForecastingCOVID192021]. Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period for which all forecasts were available (see Section \@ref(sensitivity-analysis) in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule [@gneiting_strictly_2007], meaning that in expectation the score is optimised by reporting a predictive distribution that is identical to the true data-generating distribution. Forecasters are therefore incentivised to report their true belief about the future. The WIS can be understood as a generalisation of the absolute error to quantile-based forecasts (also meaning that smaller values are better) and can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric [@funkAssessingPerformanceRealtime2019] we use captures how much probability mass of the forecast was above or below the true value (mapped to values between -1 and 1) and therefore represents a general tendency to over- or under-predict in relative terms. A value of -1 implies that all quantiles of the predictive distribution are below the observed value and a value of 1 that all quantiles are above the observed value. Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \@ref(tab:scoring-metrics) in the SI. All scores were calculated using the `scoringutils` R package [@scoringutils].

At all stages of the evaluation our forecasts were compared to the median ensemble of all *other* models submitted to the German and Polish Forecast Hub ("Hub ensemble"). This "Hub ensemble" was retrospectively computed and excludes all our models, leaving on average five ensemble member models (see details in Section \@ref(contributions-hub), as well as in Table \@ref(tab:table-ensemble-versions) and Figure \@ref(fig:num-ensemble-members) in the SI). What we call "Hub ensemble" in this article therefore differs from the "official Hub ensemble" (here called "hub-ensemble-realised") which included crowd forecasts as well as renewal model forecasts. 
To enhance interpretability of scores we mainly report WIS relative to the Hub ensemble in the main text, i.e. we divided the average scores for a given model by the average score achieved by the Hub ensemble on the same set of forecasts (with values >1 implying worse and values <1 implying better performance than the Hub ensemble). In addition to comparing our forecasts against the hub ensemble excluding our models, we also assessed the impact of our forecasts on the performance of the forecasting hub by recalculating separate versions of the Hub ensemble with only some (or all) of our forecasts included. Versions that included either all of our models (“hub-ensemble-with-all”) or only one of them (“hub-ensemble-with-X”) were computed retrospectively.

@@

```{r}
caption <- "A, C: Visualisation of 50\\% prediction intervals of two week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS."
```

@@

For cases, crowd forecasts had a lower mean weighted interval score (WIS, lower values indicate better performance) than both the renewal model and the Hub ensemble across all forecast horizons (Figure \@ref(fig:agg-performance-all)A) and locations (Figure \@ref(fig:performance-locations-rel)A). For two week ahead forecasts, mean WIS relative to the Hub ensemble (= 1) was 0.89 for crowd forecasts and 1.40 for the renewal model (Table \@ref(tab:score-table-2)). Across all forecasting approaches, locations and forecast horizons, the distribution of WIS values was very right-skewed, and average performance was heavily influenced by outliers (Figure \@ref(fig:distribution-scores)). Overall, low variance in forecast performance was closely linked with good mean performance (Figures \@ref(fig:agg-performance-all)H and and \@ref(fig:agg-performance-all)A), suggesting that the ability to avoid large errors was an important factor in determining overall performance. The impact of outlier values was especially pronounced for the renewal model, which had more outliers (Figure \@ref(fig:distribution-scores)), as well as the highest standard deviation of WIS values (standard deviation of the WIS relative to the WIS sd of the Hub ensemble was 1.54 at the two weeks ahead horizon), while the ensemble of crowd forecasts (rel. WIS sd 0.76) and the Hub ensemble (= 1) showed more stable performance.

To varying degrees, all forecasts exhibited trend-following behaviour and were rarely able to predict a change in trend before it had happened. For example, all forecasts failed to predict the change in trend from increase to decrease that happened in November in Germany and severely overshot reported cases (Figure \@ref(fig:forecasts-and-truth)A). This was most striking for the renewal model, which extrapolated unconstrained exponential growth based on the recent past of observations. The Hub ensemble and the crowd forecast, which had both been under-predicting throughout October, also failed to predict the change in trend after cases peaked, but less severely so. Human forecasters, possibly aware of the semi-lockdown announced on November 2nd 2020 [@deutschewelleCoronavirusGermanyImpose2020] and the change in the testing regime (with stricter test criteria) on November 11th 2020 [@aerzteblattSARSCoV2DiagnostikRKIPasst2020], were fastest to adapt to the new trend, and the Hub ensemble slowest. In December, cases rose again in Germany, with all models under-predicting this growth to varying extents. As in October, the renewal model captured the phase of exponential growth in cases slightly better than other approaches, but again overshot when reported case numbers fell over Christmas. The large variance in predictions in January in Germany (severe under-prediction followed by severe over-prediction) may in part be caused by the fact that the renewal model operated on daily data and therefore was susceptible to fluctuations in daily reporting around Christmas that would not have influenced on weekly reporting. Similar trends in performance were evident in Poland, with the crowd forecast quickest at adapting to the change in trend in November. In general, there were fewer large outlier forecasts in Poland and in particular the renewal model performed more in line with other forecasts there.

All forecasting approaches, including the Hub ensemble, were overconfident, i.e. they showed lower than nominal coverage (meaning that 50% (90%) prediction intervals generally covered less than 50% (90%) of the actually observed values) (Figure \@ref(fig:agg-performance-all)C and \@ref(fig:agg-performance-all)D). Coverage for all forecasts deteriorated with increasing forecast horizon, indicating that all forecasting approaches struggled to quantify uncertainty appropriately for case forecasts. This was especially an issue for crowd forecasts, which had markedly shorter prediction intervals (i.e., narrower and more confident predictive distributions) than other approaches (Figure \@ref(fig:agg-performance-all)E) and only showed a small increase in uncertainty across forecast horizons. The crowd forecasts prediction intervals were also noticeably narrower than the default baseline shown to forecasters in the application (see Figure \@ref(fig:compare-forecasters)).

In spite of good performance in terms of the absolute error (Figure \@ref(fig:agg-performance-all)G), the narrow forecast intervals led to forecasts which were severely overconfident (covering only 36% and 55% of all observations with the 50% and 90% prediction intervals of all forecasts made at a two week forecast horizon, and only 5% and 38% four weeks ahead) (Figure \@ref(fig:agg-performance-all)C,D and Tables \@ref(tab:score-table-2) and \@ref(tab:score-table-4)). Despite worse performance in terms of absolute error (Figure \@ref(fig:agg-performance-all)G), the renewal model achieved better calibration (comparable to the Hub ensemble), as uncertainty increased rapidly across forecast horizons. The crowd forecasts, on the other hand, showed a smaller bias than the renewal model, but were overconfident.

The renewal model exhibited a noticeable tendency towards over-predicting reported cases across all horizons. The crowd forecast tended to over-predict at longer forecast horizons, whereas the Hub ensemble showed no systematic bias (Figure \@ref(fig:agg-performance-all)F). Regardless of a general relative tendency to over-predict, all forecasting approaches incurred larger absolute penalties from over- than from under-prediction (see decomposition of the WIS into absolute penalties for over-prediction, under-prediction and dispersion in Figures \@ref(fig:agg-performance-all)A and \@ref(fig:agg-performance-all)B and Tables \@ref(tab:score-table-2) and \@ref(tab:score-table-4)).

Generally, trends in overall performance were broadly similar across locations (Figures \@ref(fig:performance-locations) and \@ref(fig:performance-locations-rel)).
Due to the differing population sizes and numbers of notifications in Germany and Poland absolute scores were difficult to compare directly. However, relative to the Hub ensemble, the crowd forecasts performed noticeably better in Germany than in Poland and the renewal model better in Poland than in Germany (Figures \@ref(fig:performance-locations-rel)A, \@ref(fig:performance-locations-rel)G, \@ref(fig:agg-performance-all-Germany), \@ref(fig:agg-performance-all-Poland)).

@@

For deaths, the Hub ensemble outperformed the crowd forecasts as well as our model-based approaches across all forecast horizons and locations (Figure \@ref(fig:agg-performance-all)B, Figure \@ref(fig:performance-locations)B). Relative WIS values for the models two weeks ahead were 1.22 (convolution model), 1.26 (crowd forecast), 1 (Hub ensemble) and 1.79 (renewal model). The crowd forecasts performed better than the renewal model across all forecast horizons and locations (Figure \@ref(fig:agg-performance-all)B, Figure \@ref(fig:performance-locations)B), and also better than the convolution model three and four weeks ahead. Poor performance of the renewal model, especially at longer horizons, indicates that an approach that does not know about past cases, but instead estimates and projects a separate $R_t$ trace from deaths, does not use the available information efficiently. The convolution model was able to outperform both the renewal model and the crowd forecasts at shorter forecast horizons (where the delay between cases and deaths means that future deaths are largely informed by present cases), but saw performance deteriorate at three and four weeks ahead (where case predictions from the renewal model were increasingly used to inform death predictions) (Figure \@ref(fig:agg-performance-all)B, Table \@ref(tab:score-table-4)).

@@

Death forecasts, generally, showed greater coverage of the 50% and 90% prediction intervals than case forecasts and no decrease in coverage across forecast horizons, indicating that it might be easier to appropriately quantify uncertainty for death forecasts. The Hub ensemble had the greatest coverage, with empirical coverage of the 50% and 90% prediction intervals exceeding 50%, and 90%, respectively, across all forecast horizons. Coverage for the crowd forecasts and our model-based approaches was generally lower than that of the Hub ensemble and mostly slightly lower than nominal coverage (Figure \@ref(fig:agg-performance-all)C and \@ref(fig:agg-performance-all)D). As for cases, the crowd forecast tended to have the narrowest prediction intervals and uncertainty increased most slowly across forecast horizons, and the renewal model forecasts generally were widest. The convolution model had relatively narrow prediction intervals for short forecast horizons, but had rapidly (and non-linearly) increasing uncertainty for longer forecast horizons, driven by increasing uncertainty in the underlying case forecasts.

>>>>

---
title: "Comparing human and model-based forecasts of COVID-19 in Germany and Poland"
header-includes:
    - \usepackage{lineno}
    - \usepackage[T1]{fontenc}
    - \linenumbers
    -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  rticles::plos_article:
    extra_dependencies: 
      caption: null
    keep_tex:  true
    toc: false
linestretch: 2 
include-header: 
  \usepackage[T1]{fontenc}
author:
  - name: Nikos I. Bosse
    email: nikos.bosse@lshtm.ac.uk
    affiliation: London School of Hygiene & Tropical Medicine
    corresponding: nikos.bosse@lshtm.ac.uk
  - name: Sam Abbott
    email: sam.abbott@lshtm.ac.uk
    affiliation: 
      - London School of Hygiene & Tropical Medicine
  - name: Johannes Bracher
    email: johannes.bracher@kit.edu
    affiliation: Karlsruhe Institute of Technology Institute of Economic Theory and Statistics
  - name: Habakuk Hain
    email: habakuk.hain@mpinat.mpg.de
    affiliation: Max Planck Institute for Multidisciplinary Sciences
  - name: Billy J. Quilty
    email: Billy.Quilty@lshtm.ac.uk
    affiliation: London School of Hygiene & Tropical Medicine
  - name: Mark Jit
    email: Mark.Jit@lshtm.ac.uk
    affiliation: London School of Hygiene & Tropical Medicine
  - name: Centre for the Mathematical Modelling of Infectious Diseases COVID-19 Working Group
    email: cmmid@lshtm.ac.uk
    affiliation: London School of Hygiene & Tropical Medicine
  - name: Edwin van Leeuwen
    email: Edwin.VanLeeuwen@phe.gov.uk
    affiliation:
      - UK Health Security Agency
      - London School of Hygiene & Tropical Medicine
  - name: Anne Cori
    email: a.cori@imperial.ac.uk
    affiliation: 
      - Imperial College London
  - name: Sebastian Funk
    email: sebastian.funk@lshtm.ac.uk
    affiliation: London School of Hygiene & Tropical Medicine
address:
  - code: London School of Hygiene & Tropical Medicine
    address: London School of Hygiene & Tropical Medicine, Keppel Street, London WC1E7HT, United Kingdom
  - code: Karlsruhe Institute of Technology Institute of Economic Theory and Statistics
    address: Karlsruhe Institute of Technology - Institute of Economic Theory and Statistics, Post Box 6980, 76049 Karlsruhe, Germany
  - code: Max Planck Institute for Multidisciplinary Sciences
    address: Max Planck Institute for Multidisciplinary Sciences, Am Faßberg 11, 37077 Göttingen, Germany
  - code: UK Health Security Agency
    address: UK Health Security Agency, London NW9 5EQ, United Kingdom
  - code: Imperial College London
    address: Imperial College London, Norfolk Place, London W2 1PG, United Kingdom

abstract: |
  Forecasts based on epidemiological modelling have played an important role in shaping public policy throughout the COVID-19 pandemic. This modelling combines knowledge about infectious disease dynamics with the subjective opinion of the researcher who develops and refines the model and often also adjusts model outputs. Developing a forecast model is difficult, resource- and time-consuming. It is therefore worth asking what modelling is able to add beyond the subjective opinion of the researcher alone. To investigate this, we analysed different real-time forecasts of cases of and deaths from COVID-19 in Germany and Poland over a 1-4 week horizon submitted to the German and Polish Forecast Hub. We compared crowd forecasts elicited from researchers and volunteers, against a) forecasts from two semi-mechanistic models based on common epidemiological assumptions and b) the ensemble of all other models submitted to the Forecast Hub. We found crowd forecasts, despite being overconfident, to outperform all other methods across all forecast horizons when forecasting cases (weighted interval score relative to the Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational models performed comparably better when predicting deaths (rel. WIS 1.26), suggesting that epidemiological modelling and human judgement can complement each other in important ways.

author_summary: |
  Mathematical models of COVID-19 have played a key role in informing governments across the world. While mathematical models are informed by our knowledge of infectious disease dynamics, they are ultimately developed and iteratively adjusted by the researchers and shaped by their subjective opinions. To investigate what modelling is able to add beyond the subjective opinion of the researcher alone, we compared human forecasts with model-based predictions of COVID-19 cases and deaths submitted to the so-called German/Polish Forecast Hub (which collates a variety of models from a range of teams). |
  We found that our human forecasts consistently outperformed an aggregate of all available model-based forecasts when predicting cases, but not when predicting deaths. Our findings suggest that human insight may be most valuable when forecasting highly uncertain quantities, which depend on many factors that are hard to model using equations, while mathematical models may be most useful in settings like predicting deaths, where leading indicators with a clear connection to the target variable are available. This potentially has very relevant policy implications, as agencies informing policy-makers could benefit from routinely eliciting human forecasts in addition to model-based predictions to inform policies.

bibliography:
  - references.bib
  - GermanPolishpaper.bib
csl: plos.csl
---

@@

@@

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish Forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control [@ecdcDownloadHistoricalData2020]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@rkiRKICoronavirusSARSCoV22021]. These data are subject to reporting artefacts, (such as for example delayed case reporting in Poland on the 24th November, @forsal.plRozbieznosciStatystykachKoronawirusa2020), changes in reporting over time, and variation in testing regimes (for example in Germany from the 11th of November on, @aerzteblattSARSCoV2DiagnostikRKIPasst2020). The ECDC data as well as the data published by the Polish Ministry of Health were also subject to data revisions, although most of them (with a notable exception of a data update for October 12 2020 in Germany) only affected daily, not weekly data (see Figures \ref{fig:daily-truth-update} and \ref{fig:weekly-truth-update}).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the `shiny` and `golem` R packages [@golem; @shiny] and is available in the `crowdforecastr` R package [@crowdforecastr]. To make a forecast in the application participants could select a predictive distribution (with the default being log-normal) to represent the probability that the forecasted quantity took certain values. Median and width of the uncertainty could be adjusted by either interacting with a figure showing their forecast or providing numerical values (see screenshot in Figure \ref{fig:screenshot} in the SI). The default shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e. as $\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )$). A comparison of the crowd forecasts against the default baseline shown in the application is displayed in Figure \ref{fig:compare-forecasters} in the Appendix. 
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from  @ourworldindataCOVID19DataExplorer2020. These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time. From November 26 2020 on we displayed weekly small reports with a visualisation of past forecasts and scores on our website, epiforecasts.io.

@@

The study was approved by the Observational / Interventions Research Ethics Committee at the London School of Hygiene & Tropical Medicine, LSHTM Ethics Reference: 22290. Formal consent from participants was taken in written form through the web application.

@@

For the main analysis we focused mostly on two week ahead forecasts, as COVID-19 forecasts, especially for cases, were in the past found to have poor predictive performance  beyond this horizon [@bracherShorttermForecastingCOVID192021]. Forecasts for cases were scored using the full period from October 2020 until March 2021. To ensure comparability between models, all death forecasts were scored using only the period from December 14th on, where all models including the convolution model were available. To ensure robustness of our results we conducted a sensitivity analysis where all forecasts (including cases) were scored only over the later period for which all forecasts were available (see Section \ref{sensitivity-analysis} in the SI). Results remained broadly unchanged.

Forecasts were analysed using the following scoring metrics: The weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], the absolute error, relative bias, and empirical coverage of the 50% and 90% prediction intervals. The WIS is a proper scoring rule [@gneiting_strictly_2007], meaning that in expectation the score is optimised by reporting a predictive distribution that is identical to the true data-generating distribution. Forecasters are therefore incentivised to report their true belief about the future. The WIS can be understood as a generalisation of the absolute error to quantile-based forecasts (also meaning that smaller values are better) and can be decomposed into three separate penalties: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction. While the over- and under-prediction components of the WIS capture the amount of over-prediction and under-prediction in absolute terms, we also look at a relative tendency to make biased forecasts. The bias metric [@funkAssessingPerformanceRealtime2019] we use captures how much probability mass of the forecast was above or below the true value (mapped to values between -1 and 1) and therefore represents a general tendency to over- or under-predict in relative terms. A value of -1 implies that all quantiles of the predictive distribution are below the observed value and a value of 1 that all quantiles are above the observed value. Empirical coverage is the percentage of observed values that fall inside a given prediction interval (e.g. how many observed values fall inside all 50% prediction intervals). Scoring metrics are explained in more detail in Table \ref{tab:scoring-metrics} in the SI. All scores were calculated using the `scoringutils` R package [@scoringutils].

At all stages of the evaluation our forecasts were compared to the median ensemble of all *other* models submitted to the German and Polish Forecast Hub ("Hub ensemble"). This "Hub ensemble" was retrospectively computed and excludes all our models, leaving on average five ensemble member models (see details in Section \ref{contributions-hub}, as well as in Table \ref{tab:table-ensemble-versions} and Figure \ref{fig:num-ensemble-members} in the SI). What we call "Hub ensemble" in this article therefore differs from the "official Hub ensemble" (here called "hub-ensemble-realised") which included crowd forecasts as well as renewal model forecasts. 
To enhance interpretability of scores we mainly report WIS relative to the Hub ensemble in the main text, i.e. we divided the average scores for a given model by the average score achieved by the Hub ensemble on the same set of forecasts (with values >1 implying worse and values <1 implying better performance than the Hub ensemble). In addition to comparing our forecasts against the hub ensemble excluding our models, we also assessed the impact of our forecasts on the performance of the forecasting hub by recalculating separate versions of the Hub ensemble with only some (or all) of our forecasts included. Versions that included either all of our models (“hub-ensemble-with-all”) or only one of them (“hub-ensemble-with-X”) were computed retrospectively.

@@

```{r}
caption <- "Two week ahead forecasts and corresponding scores. A, C: Visualisation of 50\\% prediction intervals of two week ahead forecasts against the reported values. Forecasts that were not scored (because there was no complete set of death forecasts available) are greyed out. B, D: Visualisation of corresponding WIS."
```

@@

For cases, crowd forecasts had a lower mean weighted interval score (WIS, lower values indicate better performance) than both the renewal model and the Hub ensemble across all forecast horizons (Figure \ref{fig:agg-performance-all}A) and locations (Figure \ref{fig:performance-locations-rel}A). For two week ahead forecasts, mean WIS relative to the Hub ensemble (= 1) was 0.89 for crowd forecasts and 1.40 for the renewal model (Table \ref{tab:score-table-2}). Across all forecasting approaches, locations and forecast horizons, the distribution of WIS values was very right-skewed, and average performance was heavily influenced by outliers (Figure \ref{fig:distribution-scores}). Overall, low variance in forecast performance was closely linked with good mean performance (Figures \ref{fig:agg-performance-all}H and and \ref{fig:agg-performance-all}A), suggesting that the ability to avoid large errors was an important factor in determining overall performance. The impact of outlier values was especially pronounced for the renewal model, which had more outliers (Figure \ref{fig:distribution-scores}), as well as the highest standard deviation of WIS values (standard deviation of the WIS relative to the WIS sd of the Hub ensemble was 1.54 at the two weeks ahead horizon), while the ensemble of crowd forecasts (rel. WIS sd 0.76) and the Hub ensemble (= 1) showed more stable performance.

To varying degrees, all forecasts exhibited trend-following behaviour and were rarely able to predict a change in trend before it had happened. For example, all forecasts failed to predict the change in trend from increase to decrease that happened in November in Germany and severely overshot reported cases (Figure \ref{fig:forecasts-and-truth}A). This was most striking for the renewal model, which extrapolated unconstrained exponential growth based on the recent past of observations. The Hub ensemble and the crowd forecast, which had both been under-predicting throughout October, also failed to predict the change in trend after cases peaked, but less severely so. Human forecasters, possibly aware of the semi-lockdown announced on November 2nd 2020 [@deutschewelleCoronavirusGermanyImpose2020] and the change in the testing regime (with stricter test criteria) on November 11th 2020 [@aerzteblattSARSCoV2DiagnostikRKIPasst2020], were fastest to adapt to the new trend, and the Hub ensemble slowest. In December, cases rose again in Germany, with all models under-predicting this growth to varying extents. As in October, the renewal model captured the phase of exponential growth in cases slightly better than other approaches, but again overshot when reported case numbers fell over Christmas. The large variance in predictions in January in Germany (severe under-prediction followed by severe over-prediction) may in part be caused by the fact that the renewal model operated on daily data and therefore was susceptible to fluctuations in daily reporting around Christmas that would not have influenced on weekly reporting. Similar trends in performance were evident in Poland, with the crowd forecast quickest at adapting to the change in trend in November. In general, there were fewer large outlier forecasts in Poland and in particular the renewal model performed more in line with other forecasts there.

All forecasting approaches, including the Hub ensemble, were overconfident, i.e. they showed lower than nominal coverage (meaning that 50% (90%) prediction intervals generally covered less than 50% (90%) of the actually observed values) (Figure \ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D). Coverage for all forecasts deteriorated with increasing forecast horizon, indicating that all forecasting approaches struggled to quantify uncertainty appropriately for case forecasts. This was especially an issue for crowd forecasts, which had markedly shorter prediction intervals (i.e., narrower and more confident predictive distributions) than other approaches (Figure \ref{fig:agg-performance-all}E) and only showed a small increase in uncertainty across forecast horizons. The crowd forecasts prediction intervals were also noticeably narrower than the default baseline shown to forecasters in the application (see Figure \ref{fig:compare-forecasters}).

In spite of good performance in terms of the absolute error (Figure \ref{fig:agg-performance-all}G), the narrow forecast intervals led to forecasts which were severely overconfident (covering only 36% and 55% of all observations with the 50% and 90% prediction intervals of all forecasts made at a two week forecast horizon, and only 5% and 38% four weeks ahead) (Figure \ref{fig:agg-performance-all}C,D and Tables \ref{tab:score-table-2} and \ref{tab:score-table-4}). Despite worse performance in terms of absolute error (Figure \ref{fig:agg-performance-all}G), the renewal model achieved better calibration (comparable to the Hub ensemble), as uncertainty increased rapidly across forecast horizons. The crowd forecasts, on the other hand, showed a smaller bias than the renewal model, but were overconfident.

The renewal model exhibited a noticeable tendency towards over-predicting reported cases across all horizons. The crowd forecast tended to over-predict at longer forecast horizons, whereas the Hub ensemble showed no systematic bias (Figure \ref{fig:agg-performance-all}F). Regardless of a general relative tendency to over-predict, all forecasting approaches incurred larger absolute penalties from over- than from under-prediction (see decomposition of the WIS into absolute penalties for over-prediction, under-prediction and dispersion in Figures \ref{fig:agg-performance-all}A and \ref{fig:agg-performance-all}B and Tables \ref{tab:score-table-2) and \ref{tab:score-table-4}).

Generally, trends in overall performance were broadly similar across locations (Figures \ref{fig:performance-locations) and \ref{fig:performance-locations-rel}).
Due to the differing population sizes and numbers of notifications in Germany and Poland absolute scores were difficult to compare directly. However, relative to the Hub ensemble, the crowd forecasts performed noticeably better in Germany than in Poland and the renewal model better in Poland than in Germany (Figures \ref{fig:performance-locations-rel}A, \ref{fig:performance-locations-rel}G, \ref{fig:agg-performance-all-Germany}, \ref{fig:agg-performance-all-Poland}).

@@

For deaths, the Hub ensemble outperformed the crowd forecasts as well as our model-based approaches across all forecast horizons and locations (Figure \ref{fig:agg-performance-all}B, Figure \ref{fig:performance-locations}B). Relative WIS values for the models two weeks ahead were 1.22 (convolution model), 1.26 (crowd forecast), 1 (Hub ensemble) and 1.79 (renewal model). The crowd forecasts performed better than the renewal model across all forecast horizons and locations (Figure \ref{fig:agg-performance-all}B, Figure \ref{fig:performance-locations}B), and also better than the convolution model three and four weeks ahead. Poor performance of the renewal model, especially at longer horizons, indicates that an approach that does not know about past cases, but instead estimates and projects a separate $R_t$ trace from deaths, does not use the available information efficiently. The convolution model was able to outperform both the renewal model and the crowd forecasts at shorter forecast horizons (where the delay between cases and deaths means that future deaths are largely informed by present cases), but saw performance deteriorate at three and four weeks ahead (where case predictions from the renewal model were increasingly used to inform death predictions) (Figure \ref{fig:agg-performance-all}B, Table \ref{tab:score-table-4}).

@@

Death forecasts, generally, showed greater coverage of the 50% and 90% prediction intervals than case forecasts and no decrease in coverage across forecast horizons, indicating that it might be easier to appropriately quantify uncertainty for death forecasts. The Hub ensemble had the greatest coverage, with empirical coverage of the 50% and 90% prediction intervals exceeding 50%, and 90%, respectively, across all forecast horizons. Coverage for the crowd forecasts and our model-based approaches was generally lower than that of the Hub ensemble and mostly slightly lower than nominal coverage (Figure \ref{fig:agg-performance-all}C and \ref{fig:agg-performance-all}D). As for cases, the crowd forecast tended to have the narrowest prediction intervals and uncertainty increased most slowly across forecast horizons, and the renewal model forecasts generally were widest. The convolution model had relatively narrow prediction intervals for short forecast horizons, but had rapidly (and non-linearly) increasing uncertainty for longer forecast horizons, driven by increasing uncertainty in the underlying case forecasts.

For deaths, the ensemble of crowd forecasts had a consistent tendency to over-predict \ref{fig:agg-performance-all}F. The convolution model had a strong tendency to under-predict, with the magnitude of under-prediction steadily decreasing for longer forecast horizons. The renewal model (which over-predicted for cases) and the Hub ensemble slightly tended towards under-prediction. For deaths, absolute over- and under-prediction penalties were more in line with a general relative tendency to over- or under-predict than for cases (Figure \ref{fig:agg-performance-all}A, \ref{fig:agg-performance-all}B and Tables \ref{tab:score-table-2}, \ref{tab:score-table-4}).

```{r}
caption <- "Distribution of scores. A: Distribution of weighted interval scores for two week ahead forecasts of the different models and forecast targets. Points denote single forecasts scores, while the shaded area shows an estimated probability density. B: Distribution of WIS separate by country. Black squares indicate median and black circles mean scores."
```
